\section{Force interpolation}

\noindent Forces are going to be computed as
with finite differences of the potential $\phi$.
It is the value of $\phi$ what we need to interpolate 
from the knowledge of the value it takes on 
the grid points.

Once $\tilde\phi$ is known for all modes on the grid,
then according to Theorem \ref{th:nyquist} 
we are able to recover $\phi(x)$ exactly 
at every point $0 \le 0 < L$ using the Fourier expansion:
\begin{equation}
    \phi(x) = \sum_{k=-k_N}^{k_N} \tilde \phi_k\, u_k(x).
\end{equation}
However, using the previous equation would imply a computational
cost of $\mathcal{O}(N)$ for the computation of the force
acting on every particle. Too expensive to compute in practice.
A method involving $\mathrm{O}(1)$ operation must be deduced
for that task, since the discrete Fourier transform 
of $\tilde\phi_k$ gives the values $\phi_n = \phi(x_n)$
of the potential at the grid points.
And for the points outside there must an interpolation
formula that computes the exact or approximate value
of $\phi$ from the knowledge of $\phi_n$ at neighboring grid points.
Different degrees of splines can be constructed for this
purpose.

The reconstruction of the function $\phi$ everywhere
from the values $\phi_n$ can be made explict.
Let's us start by evaluating the convolution $\phi * W$,
where $W$ is not yet specified:
\begin{align}
    \begin{split}
        \left( \phi* W \right)(x) 
            &= \int dx'\, \phi(x-x')\, W(x') \\
            &= \sum_{k=-\infty}^{\infty}\int dx'\,\langle u_k,\phi\rangle\,u_k(x-x')\, W(x')\\
            &= \sum_{k=-\infty}^{\infty}\int dx'\,\langle u_k,\phi\rangle\,
                u_k(x)\,u_k^{\dagger}(x')\, W(x') \\
            &= \sum_{k=-\infty}^{\infty} \,\langle u_k,\phi\rangle\,
                u_k(x)\, L\langle u_k, W\rangle. \\
    \end{split}
\end{align}
If $\phi$ satisfies the necessary conditions
for Theorem \ref{th:nyquist}, then 
we write $\langle u_k,\phi\rangle$ as a sum over the sampled values
at grid points
\begin{align}
    \begin{split}
        \left( \phi* W \right)(x) 
            &=\frac{1}{N} \sum_{n=0}^{N-1} \sum_{k=-\infty}^{\infty} \,
                \phi_n\, \omega^{-kn} \,
                u_k(x)\, L\langle u_k, W\rangle \\
            &=\frac{1}{N} \sum_{n=0}^{N-1} \sum_{k=-\infty}^{\infty} \,
                \phi_n\,
                u_k\left(x-\frac{n}{N}L\right)\, L\langle u_k, W\rangle \\
            &=\frac{L}{N} \sum_{n=0}^{N-1} \,
                \phi_n\,
                 W\left(x-\frac{n}{N}L\right). \\
    \end{split}
\end{align}
If it happens that $\phi = \phi* W$, like in the case of $\Wlow$ we
obtain:
\begin{equation}
    \phi(x) = 
            \frac{L}{N} \sum_{n=0}^{N-1} \,
                \phi_n\,
                 \Wlow\left(x-\frac{n}{N}L\right).
   \label{eq:whittaker}
\end{equation}
And that's an exact result knows as \emph{Whittaker-Shannon interpolation formula}.
Unfortunately, equation (\ref{eq:whittaker}) gives us again 
an algorithmic complexity of $\mathcal{O}(N)$ for the interpolation
since $\Wlow$ is non-local.
The literature \cite{hockney} proposes the use of the formula (\ref{eq:whittaker})
with other filters which have a compact support, like NGP, CIC, etc,
but bear in mind that we commit to an error than can be traced
back to the fact that $\phi \ne \phi * W$.

is made explicit
with the Whittaker-Shannon interpolation formula.



\noindent Ancora da vedere in \cite{hockney}.
